---
title: "Data Science Capstone Milestone Report"
author: "Connor Claypool"
date: "30 June 2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE, cache = TRUE)
```

# 1. Summary of Corpus Files

The given dataset is comprised of three English language files, each containing textual data from either blogs, news sources or Twitter. The documentation provided with the dataset states that each line of each file represents a separate entry (e.g. a sentence, paragraph, or Tweet). Table 1 provides some descriptive statistics with respect to each of these files, while Figure 1 represents these summary statistics visually. Observations based on these statistics include that the Twitter file is smaller than the others in terms of characters and words, but much longer in terms of lines. This makes sense considering the average line length for the Twitter file is much lower than for the others. Additionally, the news dataset has the longest average word length, while the Twitter dataset has the shortest.

```{r load data}

# define path and generate full filenames
path <- "~/Projects/swiftkey-nlp/data/final/en_US/"
docs <- c("blogs", "news", "twitter")
filenames <- sapply(docs, (function(f) { paste0(path, "en_US.", f, ".txt") }))
# read lines of each file
data <- sapply(filenames, readLines)

```

```{r calculate file statistics}

library(stringi)

# calculate statistics for each file
char_counts <- sapply(data, (function(c) { sum(nchar(c)) }))
word_counts <- sapply(data, (function(c) { sum(stri_stats_latex(c)[["Words"]]) }))
line_counts <- sapply(data, length)
mean_line_chars <- char_counts / line_counts
mean_line_words <- word_counts / line_counts
mean_word_length <- char_counts / word_counts

# generate data frame
file_stats <- data.frame(document = docs,
                         char.count = char_counts, 
                         word.count = word_counts, 
                         line.count = line_counts,
                         mean.line.chars = mean_line_chars,
                         mean.line.words = mean_line_words, 
                         mean.word.length = mean_word_length,
                         row.names = NULL)

```

```{r print file statistics table, results='asis', tidy=FALSE}
library(knitr)

# generate column names for table
file_stats_table_col_names <- c("Document",
                                "Character Count",
                               "Word Count",
                               "Line Count",
                               "Mean Line Length (Characters)",
                               "Mean Line Length (Words)",
                               "Mean Word Length")

# generate table
kable(file_stats, 
      caption = "Table 1: Summary Statistics Calculated for Each US English Corpus File", 
      format.args = list(big.mark=","), 
      col.names = file_stats_table_col_names)
```

```{r plot file statistics}

library(ggplot2)
library(grid)
library(gridExtra)

# barplot of character counts
char_count_plot <- ggplot(file_stats) + 
  aes(x = document, y = char.count, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Character Counts") +
  theme(legend.position="none")
# barplot of word counts
word_count_plot <- ggplot(file_stats) + 
  aes(x = document, y = word.count, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Word Counts") +
  theme(legend.position="none")
# barplot of line counts
line_count_plot <- ggplot(file_stats) + 
  aes(x = document, y = line.count, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Line Counts") +
  theme(legend.position="none")
# barplot of mean line lengths (characters)
mean_line_chars_plot <- ggplot(file_stats) + 
  aes(x = document, y = mean.line.chars, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Mean Line Lengths\n(Characters)") +
  theme(legend.position="none")
# barplot of mean line lengths (words)
mean_line_words_plot <- ggplot(file_stats) + 
  aes(x = document, y = mean.line.words, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Mean Line Lengths\n(Words)") +
  theme(legend.position="none")
# barplot of mean word lengths
mean_word_length_plot <- ggplot(file_stats) + 
  aes(x = document, y = mean.word.length, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Mean Word Lengths") +
  theme(legend.position="none")

# arrange plots into a grid
grid.arrange(char_count_plot, 
             word_count_plot, 
             line_count_plot, 
             mean_line_chars_plot, 
             mean_line_words_plot, 
             mean_word_length_plot, 
             ncol = 3, 
             bottom="Figure 1: Plots of Summary Statistics Calculated for Each US English Corpus File")
```

# 2. Data Processing and Further Exploratory Analysis

Following these basic file statistics, a more in-depth set of exploratory analyses were performed. Due to resource constraints, futher analyses were performed on a random 1% sample of the corpus. Once sampled in this way, the data was processed as follows. The sample documents were combined into a single corpus, which was then split into tokens (words), while removing punctuation, numbers and URLs. Profanity was filtered out, all tokens were converted to lowercase, and each token which appeared fewer than five times in the corpus was removed. Additionally, the tokenized corpus was used to create two more tokenized versions of the corpus, grouping tokens into 2-grams (overlapping word pairs) and 3-grams (overlapping word triplets). For each set of tokenized data, two matrices listing each unique token and its frequency were created: one with so-called stopwords included and one without.

```{r sample data}
# set seed for reproducability
set.seed(1234)

path <- "~/Projects/swiftkey-nlp/data/"

for (doc in docs) {
  # get filename
  filename <- paste0(path, "final/en_US/en_US.", doc, ".txt")
  sample_filename <- paste0(path, "sample_", doc, ".txt")
  # read data
  data <- readLines(file(filename))
  # generate logical vector of random 1% of lines to keep
  keep <- as.logical(rbinom(data, 1, 0.01))
  # subset using this vector
  sample <- data[keep]
  # write samples to new files
  writeLines(sample, file(sample_filename))
}
```

```{r process sample data}

library(readtext)
library(quanteda)
library(dplyr)

# load sample data
path <- "~/Projects/swiftkey-nlp/"
sample_data <- readtext(paste(path, "/data/*.txt", sep=""))
# create corpus
sample_corpus <- corpus(sample_data)
# load profanity dictionary
profanity <- read.csv(paste(path, "profanity.csv", sep=""), sep=";", header = FALSE, stringsAsFactors = FALSE)$V1
# generate tokens object
sample_tokens <- sample_corpus %>% 
  tokens(remove_punct=TRUE, remove_numbers=TRUE, remove_twitter=TRUE, remove_url=TRUE) %>%
  tokens_select(profanity, selection='remove') %>%
  tokens_tolower()
# generate dfm object
sample_dfm <- dfm(sample_tokens)
# generate list of features occuring 4 or fewer times
features_le4 <- textstat_frequency(sample_dfm) %>% filter(frequency <= 4) %>% .$feature
# remove these features from tokens and dfm objects
sample_tokens <- sample_tokens %>% tokens_select(features_le4, selection='remove')
sample_dfm <- dfm(sample_tokens)
# generate 2- and 3-gram tokens objects
sample_tokens_2gram <- tokens_ngrams(sample_tokens, 2)
sample_tokens_3gram <- tokens_ngrams(sample_tokens, 3)
# generate 2- and 3-gram dfm objects
sample_dfm_2gram <- dfm(sample_tokens_2gram)
sample_dfm_3gram <- dfm(sample_tokens_3gram)
# generate feature frequency matrices for 1-, 2- and 3-grams
freq_df <- textstat_frequency(sample_dfm) %>% mutate(feature = as.factor(feature))
freq_df_2gram <- textstat_frequency(sample_dfm_2gram) %>% mutate(feature = as.factor(feature))
freq_df_3gram <- textstat_frequency(sample_dfm_3gram) %>% mutate(feature = as.factor(feature))


# create token and dfm objects as above but with stopwords removed
sample_tokens_ns <- sample_tokens %>% tokens_select(stopwords("english"), selection = 'remove')
sample_dfm_ns <- dfm(sample_tokens_ns)

sample_tokens_2gram_ns <- tokens_ngrams(sample_tokens_ns, 2)
sample_tokens_3gram_ns <- tokens_ngrams(sample_tokens_ns, 3)
sample_dfm_2gram_ns <- dfm(sample_tokens_2gram_ns)
sample_dfm_3gram_ns <- dfm(sample_tokens_3gram_ns)

freq_df_ns <- textstat_frequency(sample_dfm_ns) %>% mutate(feature = as.factor(feature))
freq_df_2gram_ns <- textstat_frequency(sample_dfm_2gram_ns) %>% mutate(feature = as.factor(feature))
freq_df_3gram_ns <- textstat_frequency(sample_dfm_3gram_ns) %>% mutate(feature = as.factor(feature))
```

One basic impact of removing stopwords can be seen in Table 2. Removing this relatively small number of unique words resulted in a large decrease in the total word count. Clearly, stopwords are used incredibly frequently, despite having relatively little importance, by definition, in terms of adding meaning to the English language.

```{r feature counts}

# calculate total tokens and features w/ and w/o stopwords
total_tokens <- sum(ntoken(sample_dfm))
total_features <- nfeat(sample_dfm)
total_tokens_ns <- sum(ntoken(sample_dfm_ns))
total_features_ns <- nfeat(sample_dfm_ns)

# generate and display table of total tokens and features w/ and w/o stopwords
fcount <- data.frame(stopwords.included = c("Yes", "No"), 
                     totals = c(total_tokens, total_tokens_ns), 
                     unique = c(total_features, total_features_ns))

kable(fcount, 
      col.names = c("Stopwords Included", "Total Tokens", "Unique Tokens"),
      format.args = list(big.mark=","),
      caption = "Table 2: Total and Unique Token Counts")

```

## 2.1. N-Gram Token Frequencies (Stopwords Removed)

Figure 2 shows a wordcloud generated from the 1-gram no-stopwords frequency matrix. Figures 3, 4 and 5 display the top 30 no-stopwords 1-, 2- and 3-grams by frequency, respectively.

```{r wordcloud stopwords removed}

textplot_wordcloud(sample_dfm_ns, max_words = 100)

```

Figure 2: Wordcloud (Stopwords Removed)

```{r exploratory graphs stopwords removed}

# get top 30 1-, 2- and 3-grams by frequency (no stopwords)
freq_top30_ns <- freq_df_ns %>% slice(1:30)
freq_2gram_top30_ns <- freq_df_2gram_ns %>% slice(1:30)
freq_3gram_top30_ns <- freq_df_3gram_ns %>% slice(1:30)

# plot top 30 1-grams
ggplot(freq_top30_ns) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 3: Top 30 1-Grams by Frequency (Stopwords Removed)") +
  xlab("feature (1-gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 2-grams
ggplot(freq_2gram_top30_ns) + 
  aes(x = reorder(feature, frequency), y = frequency) +
  ggtitle("Figure 4: Top 30 2-Grams by Frequency (Stopwords Removed)") +
  xlab("feature (2-gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 3-grams
ggplot(freq_3gram_top30_ns) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 5: Top 30 3-Grams by Frequency (Stopwords Removed)") +
  xlab("feature (3-gram)") +
  geom_col(fill='darkred') + 
  coord_flip()
```

## 2.2. N-Gram Token Frequencies (Stopwords Included)

Figure 6 shows a wordcloud generated from the 1-gram stopword-inclusive frequency matrix. Figures 7, 8 and 9 display the top 30 stopword-inclusive 1-, 2- and 3-grams by frequency, respectively. It can be seen from these figures that stopwords are indeed used very frequently.

```{r wordcloud stopwords included}

textplot_wordcloud(sample_dfm, max_words = 100, min_size=1)

```

Figure 6: Wordcloud (Stopwords Included)

```{r exploratory graphs stopwords included}

# get top 30 1-, 2- and 3-grams by frequency (stopwords included)
freq_top30 <- freq_df %>% slice(1:30)
freq_2gram_top30 <- freq_df_2gram %>% slice(1:30)
freq_3gram_top30 <- freq_df_3gram %>% slice(1:30)

# plot top 30 1-grams
ggplot(freq_top30) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 7: Top 30 1-grams by Frequency (Stopwords Included)") +
  xlab("feature (1-Gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 2-grams
ggplot(freq_2gram_top30) + 
  aes(x = reorder(feature, frequency), y = frequency) +
  ggtitle("Figure 8: Top 30 2-Grams by Frequency (Stopwords Included)") +
  xlab("Feature (2-gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 3-grams
ggplot(freq_3gram_top30) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 9: Top 30 3-Grams by Frequency (Stopwords Included)") +
  xlab("feature (3-gram)") +
  geom_col(fill='darkred') + 
  coord_flip()
```

## 2.3. Relationship Between Unique Features and Corpus Coverage Levels

Table 3 lists the proportion of unique tokens needed to cover 50%, 90%, 98% and 99% of the corpus with stopwords removed, while Figure 10 graphs the relationship between the coverage of the no-stopwords corpus and the proportion of unique tokens needed.

```{r coverage calculations no stopwords}

# calculate minimum no. of unique features required to meet required corpus coverage level
features_needed <- function(required_coverage, freqs, total)
{
  current_feature <- 0
  current_tokens <- 0
  
  # while tokens counted is less than required amount
  while (current_tokens < total * required_coverage) {
    # increment feature counter/index
    current_feature <- current_feature + 1
    # increment count of tokens covered based on frequency of current feature
    current_tokens <- current_tokens + freqs[current_feature,]$frequency
  }
  
  # return no. of features needed
  return(current_feature)
}

# generate vector of coverage proportions in 0.01 increments
coverage <- seq(0, 1, 0.01)
# calculate required features
features_proportion <- sapply(coverage, function(n) { features_needed(n, freq_df_ns, total_tokens_ns) }) / total_features_ns
# generate data frame from these
coverage_table <- data.frame(coverage, features_proportion)

# display table for 50, 90, 98, 99%
kable(coverage_table[c(51, 91, 99, 100),],
      col.names = c("Proportion of Corpus Covered", "Proportion of Unique Features Required"),
      caption = "Table 3: Proportions of Unique Featured Required for Certain Coverage Levels of the Corpus (Stopwords Removed)",
      row.names = FALSE)

# plot features needed against coverage level
ggplot(coverage_table) + 
  aes(x = coverage, y = features_proportion) + 
  geom_line() +
  xlab("Proportion of Corpus Covered") +
  ylab("Proportion of Unique Features Required") +
  ggtitle("Figure 10: Graph of Proportion of Unique Features Required\nAgainst Corpus Coverage Level (Stopwords Removed)")

```

Table 4 lists the proportion of unique tokens needed to cover 50%, 90%, 98% and 99% of the corpus with stopwords included, while Figure 11 graphs the relationship between the coverage of the stopwords-inclusive corpus and the proportion of unique tokens needed. Compared to the no-stopwords corpus, fewer unique features are required to meet each coverage level, due to the very high frequency of many stopwords.

```{r coverage calculations}

# same as previously, but with stopwords included
coverage <- seq(0, 1, 0.01)
features_proportion <- sapply(coverage, function(n) { features_needed(n, freq_df, total_tokens) }) / total_features
coverage_table <- data.frame(coverage, features_proportion)

kable(coverage_table[c(51, 91, 99, 100),],
      col.names = c("Proportion of Corpus Covered", "Proportion of Unique Features Required"),
      caption = "Table 4: Proportions of Unique Featured Required for Certain Coverage Levels of the Corpus (Stopwords Included)",
      row.names = FALSE)

ggplot(coverage_table) + 
  aes(x = coverage, y = features_proportion) + 
  geom_line() +
  xlab("Proportion of Corpus Covered") +
  ylab("Proportion of Unique Features Required") +
  ggtitle("Figure 11: Graph of Proportion of Unique Features Required\nAgainst Corpus Coverage Level (Stopwords Included)")
```

# 3. Next Steps: Predictive Modelling

The final goal for the analysis of this dataset is to build a model which can accurately predict the next word based on previous words, while minimising resource usage as the final product must function as a web app. Based on reasearch and prior experience, three methods are being considered for the modelling stage. Stupid Backoff is the simplest method, and is known to achieve high accuracy depending on the amount of data available. Kneser-Ney Smoothing is a similar but more complex method, and may yield higher accuracy. A Recurrent Neural Network (RNN)-based language model is a more advanced option which could likely achive higher accuracy than the others, but may also require more resources. Experimentation will determine which method or methods are suitable solutions for the goal in mind.

# Appendix A: Source Code Listings

```{r listing setup, include=FALSE, cache=FALSE}
knitr::opts_chunk$set(echo = TRUE, eval = FALSE)
```

## 1. Summary of Corpus Files

```{r listing load data}

# define path and generate full filenames
path <- "~/Projects/swiftkey-nlp/data/final/en_US/"
docs <- c("blogs", "news", "twitter")
filenames <- sapply(docs, (function(f) { paste0(path, "en_US.", f, ".txt") }))
# read lines of each file
data <- sapply(filenames, readLines)

```

```{r listing calculate file statistics}

library(stringi)

# calculate statistics for each file
char_counts <- sapply(data, (function(c) { sum(nchar(c)) }))
word_counts <- sapply(data, (function(c) { sum(stri_stats_latex(c)[["Words"]]) }))
line_counts <- sapply(data, length)
mean_line_chars <- char_counts / line_counts
mean_line_words <- word_counts / line_counts
mean_word_length <- char_counts / word_counts

# generate data frame
file_stats <- data.frame(document = docs,
                         char.count = char_counts, 
                         word.count = word_counts, 
                         line.count = line_counts,
                         mean.line.chars = mean_line_chars,
                         mean.line.words = mean_line_words, 
                         mean.word.length = mean_word_length,
                         row.names = NULL)

```

```{r listing print file statistics table, results='asis', tidy=FALSE}
library(knitr)

# generate column names for table
file_stats_table_col_names <- c("Document",
                                "Character Count",
                               "Word Count",
                               "Line Count",
                               "Mean Line Length (Characters)",
                               "Mean Line Length (Words)",
                               "Mean Word Length")

# generate table
kable(file_stats, 
      caption = "Table 1: Summary Statistics Calculated for Each US English Corpus File", 
      format.args = list(big.mark=","), 
      col.names = file_stats_table_col_names)
```

```{r listing plot file statistics}

library(ggplot2)
library(grid)
library(gridExtra)

# barplot of character counts
char_count_plot <- ggplot(file_stats) + 
  aes(x = document, y = char.count, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Character Counts") +
  theme(legend.position="none")
# barplot of word counts
word_count_plot <- ggplot(file_stats) + 
  aes(x = document, y = word.count, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Word Counts") +
  theme(legend.position="none")
# barplot of line counts
line_count_plot <- ggplot(file_stats) + 
  aes(x = document, y = line.count, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Line Counts") +
  theme(legend.position="none")
# barplot of mean line lengths (characters)
mean_line_chars_plot <- ggplot(file_stats) + 
  aes(x = document, y = mean.line.chars, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Mean Line Lengths\n(Characters)") +
  theme(legend.position="none")
# barplot of mean line lengths (words)
mean_line_words_plot <- ggplot(file_stats) + 
  aes(x = document, y = mean.line.words, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Mean Line Lengths\n(Words)") +
  theme(legend.position="none")
# barplot of mean word lengths
mean_word_length_plot <- ggplot(file_stats) + 
  aes(x = document, y = mean.word.length, fill = document) + 
  geom_bar(stat = "identity") +
  xlab("") + 
  ylab("") + 
  ggtitle("Mean Word Lengths") +
  theme(legend.position="none")

# arrange plots into a grid
grid.arrange(char_count_plot, 
             word_count_plot, 
             line_count_plot, 
             mean_line_chars_plot, 
             mean_line_words_plot, 
             mean_word_length_plot, 
             ncol = 3, 
             bottom="Figure 1: Plots of Summary Statistics Calculated for Each US English Corpus File")
```

## 2. Data Processing and Further Exploratory Analysis

```{r listing sample data}
# set seed for reproducability
set.seed(1234)

path <- "~/Projects/swiftkey-nlp/data/"

for (doc in docs) {
  # get filename
  filename <- paste0(path, "final/en_US/en_US.", doc, ".txt")
  sample_filename <- paste0(path, "sample_", doc, ".txt")
  # read data
  data <- readLines(file(filename))
  # generate logical vector of random 1% of lines to keep
  keep <- as.logical(rbinom(data, 1, 0.01))
  # subset using this vector
  sample <- data[keep]
  # write samples to new files
  writeLines(sample, file(sample_filename))
}
```

```{r listing process sample data}

library(readtext)
library(quanteda)
library(dplyr)

# load sample data
path <- "~/Projects/swiftkey-nlp/"
sample_data <- readtext(paste(path, "/data/*.txt", sep=""))
# create corpus
sample_corpus <- corpus(sample_data)
# load profanity dictionary
profanity <- read.csv(paste(path, "profanity.csv", sep=""), sep=";", header = FALSE, stringsAsFactors = FALSE)$V1
# generate tokens object
sample_tokens <- sample_corpus %>% 
  tokens(remove_punct=TRUE, remove_numbers=TRUE, remove_twitter=TRUE, remove_url=TRUE) %>%
  tokens_select(profanity, selection='remove') %>%
  tokens_tolower()
# generate dfm object
sample_dfm <- dfm(sample_tokens)
# generate list of features occuring 4 or fewer times
features_le4 <- textstat_frequency(sample_dfm) %>% filter(frequency <= 4) %>% .$feature
# remove these features from tokens and dfm objects
sample_tokens <- sample_tokens %>% tokens_select(features_le4, selection='remove')
sample_dfm <- dfm(sample_tokens)
# generate 2- and 3-gram tokens objects
sample_tokens_2gram <- tokens_ngrams(sample_tokens, 2)
sample_tokens_3gram <- tokens_ngrams(sample_tokens, 3)
# generate 2- and 3-gram dfm objects
sample_dfm_2gram <- dfm(sample_tokens_2gram)
sample_dfm_3gram <- dfm(sample_tokens_3gram)
# generate feature frequency matrices for 1-, 2- and 3-grams
freq_df <- textstat_frequency(sample_dfm) %>% mutate(feature = as.factor(feature))
freq_df_2gram <- textstat_frequency(sample_dfm_2gram) %>% mutate(feature = as.factor(feature))
freq_df_3gram <- textstat_frequency(sample_dfm_3gram) %>% mutate(feature = as.factor(feature))


# create token and dfm objects as above but with stopwords removed
sample_tokens_ns <- sample_tokens %>% tokens_select(stopwords("english"), selection = 'remove')
sample_dfm_ns <- dfm(sample_tokens_ns)

sample_tokens_2gram_ns <- tokens_ngrams(sample_tokens_ns, 2)
sample_tokens_3gram_ns <- tokens_ngrams(sample_tokens_ns, 3)
sample_dfm_2gram_ns <- dfm(sample_tokens_2gram_ns)
sample_dfm_3gram_ns <- dfm(sample_tokens_3gram_ns)

freq_df_ns <- textstat_frequency(sample_dfm_ns) %>% mutate(feature = as.factor(feature))
freq_df_2gram_ns <- textstat_frequency(sample_dfm_2gram_ns) %>% mutate(feature = as.factor(feature))
freq_df_3gram_ns <- textstat_frequency(sample_dfm_3gram_ns) %>% mutate(feature = as.factor(feature))
```

```{r listing feature counts}

# calculate total tokens and features w/ and w/o stopwords
total_tokens <- sum(ntoken(sample_dfm))
total_features <- nfeat(sample_dfm)
total_tokens_ns <- sum(ntoken(sample_dfm_ns))
total_features_ns <- nfeat(sample_dfm_ns)

# generate and display table of total tokens and features w/ and w/o stopwords
fcount <- data.frame(stopwords.included = c("Yes", "No"), 
                     totals = c(total_tokens, total_tokens_ns), 
                     unique = c(total_features, total_features_ns))

kable(fcount, 
      col.names = c("Stopwords Included", "Total Tokens", "Unique Tokens"),
      format.args = list(big.mark=","),
      caption = "Table 2: Total and Unique Token Counts")

```

### 2.1. N-Gram Token Frequencies (Stopwords Removed)

```{r listing wordcloud stopwords removed}

textplot_wordcloud(sample_dfm_ns, max_words = 100)

```

```{r listing exploratory graphs stopwords removed}

# get top 30 1-, 2- and 3-grams by frequency (no stopwords)
freq_top30_ns <- freq_df_ns %>% slice(1:30)
freq_2gram_top30_ns <- freq_df_2gram_ns %>% slice(1:30)
freq_3gram_top30_ns <- freq_df_3gram_ns %>% slice(1:30)

# plot top 30 1-grams
ggplot(freq_top30_ns) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 3: Top 30 1-Grams by Frequency (Stopwords Removed)") +
  xlab("feature (1-gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 2-grams
ggplot(freq_2gram_top30_ns) + 
  aes(x = reorder(feature, frequency), y = frequency) +
  ggtitle("Figure 4: Top 30 2-Grams by Frequency (Stopwords Removed)") +
  xlab("feature (2-gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 3-grams
ggplot(freq_3gram_top30_ns) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 5: Top 30 3-Grams by Frequency (Stopwords Removed)") +
  xlab("feature (3-gram)") +
  geom_col(fill='darkred') + 
  coord_flip()
```

### 2.2. N-Gram Token Frequencies (Stopwords Included)

```{r listing wordcloud stopwords included}

textplot_wordcloud(sample_dfm, max_words = 100, min_size=1)

```

```{r listing exploratory graphs stopwords included}

# get top 30 1-, 2- and 3-grams by frequency (stopwords included)
freq_top30 <- freq_df %>% slice(1:30)
freq_2gram_top30 <- freq_df_2gram %>% slice(1:30)
freq_3gram_top30 <- freq_df_3gram %>% slice(1:30)

# plot top 30 1-grams
ggplot(freq_top30) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 7: Top 30 1-grams by Frequency (Stopwords Included)") +
  xlab("feature (1-Gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 2-grams
ggplot(freq_2gram_top30) + 
  aes(x = reorder(feature, frequency), y = frequency) +
  ggtitle("Figure 8: Top 30 2-Grams by Frequency (Stopwords Included)") +
  xlab("Feature (2-gram)") +
  geom_col(fill='darkred') +
  coord_flip()

# plot top 30 3-grams
ggplot(freq_3gram_top30) + 
  aes(x = reorder(feature, frequency), y = frequency) + 
  ggtitle("Figure 9: Top 30 3-Grams by Frequency (Stopwords Included)") +
  xlab("feature (3-gram)") +
  geom_col(fill='darkred') + 
  coord_flip()
```

### 2.3. Relationship Between Unique Features and Corpus Coverage Levels

```{r listing coverage calculations no stopwords}

# calculate minimum no. of unique features required to meet required corpus coverage level
features_needed <- function(required_coverage, freqs, total)
{
  current_feature <- 0
  current_tokens <- 0
  
  # while tokens counted is less than required amount
  while (current_tokens < total * required_coverage) {
    # increment feature counter/index
    current_feature <- current_feature + 1
    # increment count of tokens covered based on frequency of current feature
    current_tokens <- current_tokens + freqs[current_feature,]$frequency
  }
  
  # return no. of features needed
  return(current_feature)
}

# generate vector of coverage proportions in 0.01 increments
coverage <- seq(0, 1, 0.01)
# calculate required features
features_proportion <- sapply(coverage, function(n) { features_needed(n, freq_df_ns, total_tokens_ns) }) / total_features_ns
# generate data frame from these
coverage_table <- data.frame(coverage, features_proportion)

# display table for 50, 90, 98, 99%
kable(coverage_table[c(51, 91, 99, 100),],
      col.names = c("Proportion of Corpus Covered", "Proportion of Unique Features Required"),
      caption = "Table 3: Proportions of Unique Featured Required for Certain Coverage Levels of the Corpus (Stopwords Removed)",
      row.names = FALSE)

# plot features needed against coverage level
ggplot(coverage_table) + 
  aes(x = coverage, y = features_proportion) + 
  geom_line() +
  xlab("Proportion of Corpus Covered") +
  ylab("Proportion of Unique Features Required") +
  ggtitle("Figure 10: Graph of Proportion of Unique Features Required\nAgainst Corpus Coverage Level (Stopwords Removed)")

```

```{r listing coverage calculations}

# same as previously, but with stopwords included
coverage <- seq(0, 1, 0.01)
features_proportion <- sapply(coverage, function(n) { features_needed(n, freq_df, total_tokens) }) / total_features
coverage_table <- data.frame(coverage, features_proportion)

kable(coverage_table[c(51, 91, 99, 100),],
      col.names = c("Proportion of Corpus Covered", "Proportion of Unique Features Required"),
      caption = "Table 4: Proportions of Unique Featured Required for Certain Coverage Levels of the Corpus (Stopwords Included)",
      row.names = FALSE)

ggplot(coverage_table) + 
  aes(x = coverage, y = features_proportion) + 
  geom_line() +
  xlab("Proportion of Corpus Covered") +
  ylab("Proportion of Unique Features Required") +
  ggtitle("Figure 11: Graph of Proportion of Unique Features Required\nAgainst Corpus Coverage Level (Stopwords Included)")
```